---
title: "Functions of TCIU Analytics"
subtitle: "Validation using 4D Complex (mag,phase) and Real (solely-mag reconstruction) fMRI data"
author: "<h3> SOCR Team </h3>"
date: "`r format(Sys.time(),'%m/%d/%Y')`"
output: 
  html_document:
    theme: spacelab
    highlight: tango
    includes:
      before_body: TCIU_header.html
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Functions of TCIU Analytics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Background

This document is set up for package TCIU which is developed in the [R environment](https://www.r-project.org). In this vignette, we introduce the detailed usage of all the functions with examples in the TCIU R package. Besides, users can also refer to the "workflow" vignette which provides more concise steps on how to do the analyze fMRI data or the website [TCIU Predictive Analytics](https://www.socr.umich.edu/TCIU/HTMLs/TCIU_Predictive_Analytics.html).

```{r warning=FALSE, message=FALSE}
require(TCIU) 
require(DT)
require(AnalyzeFMRI)
```

# Function list

## FMRI data simulation

- **fmri_simulate_func**: an real-valued fMRI data simulation function, used to simply generate a 3 dimension area with activated parts inside.

## Visualization of the fMRI data in time-series

- **fmri_time_series**: create four interactive time series graphs for the Real, Imaginary, Magnitude, and Phase parts for the fMRI spacetime data.

- **fmri_kimesurface**: transform the fMRI time-series data at a fixed voxel location into a kime-surface.

- **fmri_image**: create images for front view, side view, and top view of the fMRI image.

- **fmri_ts_forecast**: forecast with time series of different voxels in fMRI time-series data.

## Activated areas detection

- **fmri_stimulus_detect**: takes in real/complex valued fmridata and detects locations where stimulus occurs (i.e. generate p-values).

- **fmri_post_hoc**: conduct the post-hoc process(i.e. fdr correction and spatial clustering) for a 3-dimensional p-value array.

## Activated areas visualization

- **fmri_2dvisual**: a visualization method, use ggplot2 to draw the brain from axial, sagittal and coronal view with activated area identified by p value.

- **fmri_pval_comparison_2d**: a plots arrangement method, which uses **gridExtra** to combine multiple 2d plots of the fMRI data together. This can bring convenience for users to compare the performance of different statistical tests based on the p values they generate.

- **fmri_3dvisual**: a visualization method, use **plotly** to draw the 3d plot of the brain with the activated areas determined by p value.

- **fmri_pval_comparison_3d**: a visualization method, use **plotly** to compare the activated parts inside in the brain using two sets of color palette. The activated parts are decided by p values.


# fMRI data simulation

## Funtion "fmri_simulate_func" 

To begin with the analytic, we need to have the 4D real-valued(solely-mag reconstruction) fMRI data or the 4D complex-value(mag,phase) fMRI data.

If the user doesn't have the raw data, we have provided a simulation function in the package to simulate the 4D real-valued fMRI data. We also provide a sample brain-shaped mask in the package.

**Details/Parameters**

The function fmri_simulate_func is used to simulate fMRI data with specified dimension and total time points. 
The fMRI data can own a brain shape by using the mask data provided in our package, if the dimension fits the same as our data (c(64, 64. 40)).
Otherwise, the function will generate a 3d sphere data with multiple activated part inside. The activated parts can be detected based on the p values.

Parameters in this function:

- *dim_data* : a vector of length three to identify the dimension of fMRI data wanted to simulate

- *mask* : a parameter to have the fMRI data with the brain-shaped mask provided in the package, or the 3d sphere mask generated by the function itself.

- *ons* : a vector of the start time points of the time period when the fMRI data receives stimulation.

- *dur* : a vector of the time period when the fMRI data receives stimulation.

**Simulation**
```{r warning=FALSE, message=FALSE}
# simulate the fmridata
fmri_generate = fmri_simulate_func(dim_data = c(64, 64, 40), mask = mask, 
                                   ons = c(1, 21, 41, 61, 81, 101, 121, 141), 
                                   dur = c(10, 10, 10, 10, 10, 10, 10, 10))
dim(fmri_generate$fmri_data)
```

# Visualization of the fMRI data in time-series


## Funtion "fmri_time_series" 

**Details/Parameters**

We first provide an interactive time-series visualization for a fixed voxel. Since the input fMRI data needs to be complex-valued, we provide a sample complex voxel's time-series vector and a sample reference **plotly** object.

The function fmri_time_series is used to create four interactive time series graphs for the real, imaginary, magnitude, and phase parts for the fMRI spacetime data.

Parameters in this function:

- *fmridata* : a 4d array which contains the spatial and temporal record of fMRI result or a single complex valued vector.

- *voxel_location* : By default is TRUE. If change to false, need to input a vector instead of a 4d array.

- *is.4d* : By default is TRUE. If change to false, need to input a vector instead of a 4d array.

- *ref* : By default is NULL. Outside extra reference **plotly** object that you wish to include in the final result.

**Example**

```{r warning=FALSE, message=FALSE}
sample_voxel = system.file("extdata", "sample_voxel.rda", package = "TCIU", mustWork=TRUE)
reference_plot = system.file("extdata", "reference_plot.rda", package = "TCIU", mustWork=TRUE)
load(sample_voxel)
load(reference_plot)
fmri_time_series(sample_voxel, voxel_location = NULL, is.4d = FALSE, ref = reference_plot)
```

## Funtion "fmri_kimesurface" 

**Details/Parameters**

We second provide a function to transform the fMRI time-series at a fixed voxel location into a kime-surface (kime-series).

The function fmri_kimesurface is display in 3D the kime-series as 2D manifolds (kime-surface) over the Cartesian domain. User can choose to provide the 4D array of the fMRI spacetime image and the voxel_location or a single TS vector.

Parameters in this function:

- *fmridata* : a 4d array which contains the spatial and temporal record of fMRI result or a single complex valued vector.

- *voxel_location* : By default is TRUE. If change to false, need to input a vector instead of a 4d array.

- *is.4d* : By default is TRUE. If change to false, need to input a vector instead of a 4d array.

**Example**

```{r warning=FALSE, message=FALSE}
 # a data-frame with 160 rows and 4 columns: time (1:10), phases (8), states (2), and fMRI_value (Complex or Real intensity)
 datatable(fmri_kimesurface(fmri_generate$fmri_data, c(44,30,33))[[1]])
 # ON Kime-Surface
 fmri_kimesurface(fmri_generate$fmri_data, c(44,30,33))[[2]]
 # OFF Kime-Surface 
 fmri_kimesurface(fmri_generate$fmri_data, c(44,30,33))[[3]]
 # ON&OFF Kime-Surface 
 fmri_kimesurface(fmri_generate$fmri_data, c(44,30,33))[[4]]
```

## Funtion "fmri_image" 

**Details/Parameters**

The function fmri_image is used to create images for front view, side view, and top view of the fMRI image. When providing the 4D array of the fMRI spacetime image and input the x,y,z position of the voxel, three views of the fMRI image and the time series image of the voxel will be shown.

Parameters in this function:

- *fmridata* : a 4D array contains information for the fMRI spacetime image. The data should only contains the magnitude for the fMRI image.

- *option* : The default is 'manually'. If choose 'auto', then this function will lead you to key in the space (x,y,z) parameters and time (time) parameter for this function to generate graphs.

- *voxel_location* : a 3d array indicating the spatial location of the brain. If option is auto, set the voxel_location as NULL.

- *time* : time location for the z axis

**Example**

```{r warning=FALSE, message=FALSE}
fmri_image(fmri_generate$fmri_data, option="manually", voxel_location = c(40,22,33), time=4)
```

## Funtion "fmri_ts_forecast" 

**Details/Parameters**

The function fmri_ts_forecast is used to used to forecast with time series.

Parameters in this function:

- *fmridata* : a 4D array contains information for the fMRI spacetime image. The data should only contains the magnitude for the fMRI image.

- *voxel_location* : a 3d array indicating the voxel location of the brain

- *cut* : breaking point of the time-series data. The default is 10. 

**Example**

```{r warning=FALSE, message=FALSE}
smoothmod<-GaussSmoothArray(fmri_generate$fmri_data, sigma = diag(3,3))
fmri_ts_forecast(smoothmod, voxel_location=c(41,44,33))
```

# Activated areas detection

## Funtion "fmri_stimulus_detect" 

**Details/Parameters**

The function fmri_stimulus_detect is used to conduct motor area detection. It first takes in a real or complex valued fmridata, and then users can choose to use various methods to find the spacial regions where motor area is located inside the brain. User can either input the 4d fmridata ans get a 3d array storing p-values or input the fmridata with smaller dimension(e.x. fix the x,y axis) and get a vector storing p-values. Besides, one can use this function to just calculate raw p-values, and we also provide options so that users can do fdr correction and spatial clustering to get a more accurate result.

Parameters in this function:

- *fmridata* : an array or a vector which contains the spatial and/or temporal record of fmri result

- *mask* : a 3d array indicating the spacial location of the brain

- *stimulus_idx* : a vector that specifies when motion happens 

- *rest_idex* : a vector that specifies when study participant does not move

- *method* : a string that indicates which testing method is to be used. There are 5 options: 'HotellingT2', 'Wilk's Lambda' and 'gLRT'(likelihood ratio test) for complex fmridata and 't-test', 'wilcox-test' for real fmridata. For 4D real-valued fMRI data, two more options: 'on_off_diff' and 'HRF' method can be applied.

- *fdr_corr* : a logical variable. True if fdr correction is to be applied

- *spatial_cluster.thr* : threshold p-value to be used for spatial clustering

- *spatial_cluster.size* : number of spatially connected voxels to be tested for spatial clustering

**Example**

```{r, eval=FALSE}
# If the dimension of fmridata is lower than 4d, the corresponding dimension of output p-values will be lower.
# Users can also choose other test method or apply fdr correction and spatial clustering to get the result.
# Here, we provide two examples of t-test and wilcox-test for the 4d real-valued fmri data.
p_simulate_t_test = fmri_stimulus_detect(fmridata= fmri_generate$fmri_data, 
                                         mask = fmri_generate$mask,
                                         stimulus_idx = fmri_generate$on_time,
                                         method = "t-test")
p_simulate_wilcox = fmri_stimulus_detect(fmridata= fmri_generate$fmri_data, 
                                         mask = fmri_generate$mask,
                                         stimulus_idx = fmri_generate$on_time,
                                         method = "wilcox-test")
```
## Funtion "fmri_post_hoc"

**Details/Parameters**

The function fmri_post_hoc would help do the fdr correction and spatial clustering for a 3d p-value array. The fdr correction controls for a low proportion of false positives, while the spatial clustering part help filter out all sparse p-values that are not in a specified clusters.
**Notice: The correction for 3d p-value can be directly set in stimulus detect function. So we seldom apply this function by itself.**

Parameters in this function:

- *p_val_3d* : an array which contains the p-values as the result of fMRI statistical tests.

- *fdr_corr* : The default is NULL. Input "fdr" to conduct fdr correction.

- *spatial_cluster.thr* : The default is NULL. Together with spatial_cluster.size are used to filter contiguous clusters of locations in a 3D array that are below some threshold and with some minimum size.

- *spatial_cluster.size* : The default is NULL. The size of spatial cluster.

- *show_comparison* : The default is FALSE. If TRUE, would display the comparison between raw and processed p-values.

**Example**

```{r}
# load 3D p value stored in the package
dim(pval1)
# do the fdr correction
pval_fdr = fmri_post_hoc(pval1 , fdr_corr = "fdr",
                         spatial_cluster.thr = NULL,
                         spatial_cluster.size = NULL, 
                         show_comparison = FALSE)
# do the spatial clustering
pval_posthoc = fmri_post_hoc(pval_fdr, fdr_corr = NULL,
                             spatial_cluster.thr = 0.05,
                             spatial_cluster.size = 5, 
                             show_comparison = FALSE)
```

# Activated areas visualization

A more detailed output and manipulation about this part is introduced in the "workflow" vignette.

## Funtion "fmri_2dvisual"

**Details/Parameters**

The function fmri_2dvisual is used to find activated part of the brain based on given p values from sagittal, axial and coronal view.

Parameters in this function:

- *pval* : a 3d array p value that can be used to plot activated area of the brain

- *axis_ls* :  a list with two elements. The first element is the character of x, y, z; the second element is a numeric integer to show the specific slice on the fixed axis identified in the first element

- *hemody_data* :  a parameter to have the plot with/without hemodynamic contour. The default is NULL to make the plot without hemodynamic contour, otherwise assign a 3D array of the hemodynamic data.

- *mask* : a 3d nifti or 3d array data to show the position of the brain

- *p_threshold* : a numeric value that can be selected randomly below 0.05 to drop insignificant p value of no need. If "low5_percent" method is used, make 'p_threshold' as NULL. Default is 0.05.

- *legend_show* : a logical parameter of True and False to specify whether the final plot has the legend

- *method* : a string that represent method to draw the plot. There are 3 options: "min_max", "scale_p" and "low5_percent".

- *color_pal* : name of the color palettes provided by **RColorBrewer**. The default is "YlOrRd".

- *multip_pranges* : an option under 'scale_p' method for user to decide whether there are at most 9 colors in the legend for the ranges of p value.

**Example**

```{r}
fmri_2dvisual(pval1, list('x',40), hemody_data=NULL, mask=mask, p_threshold=0.05)
fmri_2dvisual(pval1, list('y',23), hemody_data=NULL, mask=mask, p_threshold=0.05)
fmri_2dvisual(pval1, list('z',33), hemody_data=NULL, mask=mask, p_threshold=0.05)
```

## Funtion "fmri_pval_comparison_2d"

**Details/Parameters**

The function fmri_pval_comparison_2d is used to combine the 2d plots of different view generated by different 3d p value data for comparison convenience.

Parameters in this function:

- *pval_ls* : a list with each element as a 3d array of p values generated by the statistical test 

- *pval_name_ls* : a list with each element as the name for the p values provided in 'pval_ls'

- *axis_i_lses* : a list with 3 numeric elements for the specific locations of the three 2d planes of the brain, or it is a list of lists. Each sub-list also has 3 numeric elements. If the view from any axis does not want to be shown, just input the NULL for that element.

- *hemody_data* :  a parameter to have the plot with/without hemodynamic contour. The default is NULL to make the plot without hemodynamic contour, otherwise assign a 3D array of the hemodynamic data.

- *mask* : a 3d nifti or 3d array data to show the position of the brain

- *p_threshold* : a numeric value that can be selected randomly below 0.05 to drop insignificant p value of no need. If "low5_percent" method is used, make 'p_threshold' as NULL. Default is 0.05.

- *legend_show* : a logical parameter of True and False to specify whether the final plot has the legend

- *method* : a string that represent method to draw the plot. There are 2 options: "min_max" and "scale_p". 

- *color_pal* : name of the color palettes provided by **RColorBrewer**. The default is "YlOrRd".

- *multip_pranges* : an option under 'scale_p' method for user to decide whether there are at most 9 colors in the legend for the ranges of p value

**Example**

```{r}
fmri_pval_comparison_2d(list(pval1, pval2), 
                        list('pval', 'pval2'),
                        list(list(40, 26, 33), list(40, 26, 33)), 
                        hemody_data = NULL, 
                        mask = mask, p_threshold = 0.05, 
                        legend_show = FALSE, method = 'scale_p',
                        color_pal = "YlOrRd", multi_pranges=TRUE)    
```

## Funtion "fmri_3dvisual"

**Details/Parameters**

The function fmri_3dvisual is used to visualize the 3d plot of the brain with its activated parts based on given p values.

Parameters in this function:

- *pval* : a 3d array p value that can be used to plot activated area of the brain

- *mask* : a 3d nifti or 3d array data to show the position of the brain

- *p_threshold* : a numeric value that can be selected randomly below 0.05 to drop insignificant p value of no need. If "low5_percent" method is used, make 'p_threshold' as NULL. Default is 0.05.

- *method* : a string that represent method to draw the plot. There are 2 options: "scale_p" and "low5_percent". 

- *color_pal* : name of the color palettes provided by **RColorBrewer**. The default is "YlOrRd"

- *multip_pranges* : an option under 'scale_p' method for user to decide whether there are at most 9 colors in the legend for the ranges of p value

**Example**

```{r}
fmri_3dvisual(pval1, mask, p_threshold = 0.05, method="scale_p")$plot
```

## Funtion "fmri_pval_comparison_3d"

**Details/Parameters**

The function fmri_pval_comparison_3d is used to visualize and compare the 3d plot of the activated parts in the brain in one brain shell.

Parameters in this function:

- *pval_3d_ls* : a list of two element, each element is a 3d array p value that can be used to plot activated area of the brain

- *mask* : a 3d nifti or 3d array data to show the position of the brain

- *p_threshold* :  list of two element, each element is NULL or a numeric value that can be selected randomly below 0.05 to drop insignificant p value of no need

- *method_ls* :a list of two element, each element is a string that represent method to draw the plot. There are 3 options: "scale_p" and "low5_percent". 

- *color_pal_ls* : the name of the color palettes provided by **RColorBrewer**. The default is list("YlOrRd", "YlGnBu").

- *multip_pranges* : an option under 'scale_p' method for user to decide whether there are at most 9 colors in the legend for the ranges of p value

**Example**

```{r}
fmri_pval_comparison_3d(list(pval1, pval2), mask, 
                        list(0.05, 0.05), list('scale_p', 'scale_p'), multi_pranges=FALSE)
```