---
title: "Non-Euclidean Kime Manifold"
subtitle: "[Back To Contents](Contents.html)"
author: "SOCR Team "
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  html_document:
    theme: spacelab
    highlight: tango
    includes:
      before_body: SOCR_header.html
      after_body: SOCR_footer_tracker.html
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Non-Euclidean Kime Manifold

So far, we considered spacekime as 5D Euclidean space where the 2D kime is a flat space. Does it make sense to generalize kime to non-Euclidean manifold and  what is the rationale and meaning of a curved kime manifold?

We saw earlier the geometric representation of the flat spacekime using conical shapes. Let's start with some visualization of the non-Euclidean 2D kime manifold.

```{r message=FALSE,warning=FALSE}
library(plotly)
library(ggplot2)
```
```{r}

# Define a vase-generating funciton (2D), which will be rotated to form curved kime manifold
kime.manifold <- data.frame( s=seq(0, 1.3, length.out = 200) )
kime.manifold$theta <- seq(0, 2*pi, length.out = 200)
g <- function (t) { sqrt(t^3-2.09*t^2+1.097*t) }
h <- function (t) { t }

# rendering (u==s, v==theta) parametric surfaces requires x,y,z arguments to be 2D arrays
# In out case, the three coordinates have to be 200*200 parameterized tensors/arrays
x2 <- g(kime.manifold$s) %o% cos(kime.manifold$theta)
y2 <- g(kime.manifold$s) %o% sin(kime.manifold$theta)
z2 <- h(kime.manifold$s) %o% rep(1, 200)
  
p <- plot_ly(x=~x2, y=~y2, z=~z2,   # draw vertical plane 1
        #colors = c("yellow", "white"),
        type="surface", hoverinfo="none", opacity=0.5, 
        showlegend=FALSE, showscale = FALSE,
        contours = list(
      x = list(highlight = FALSE),
      y = list(highlight = FALSE),
      z = list( highlight = TRUE, highlightcolor = "blue") )) %>%
  # add_segments(x = ~x, xend = ~x, y = ~ymin, yend = ~ymax, z = ~zmin, yend = ~zmax) %>%
  #add_trace(x=~x2, y=~y2, z=~z2,    # draw vertical plane 2 (orthogonal to 1)
  #      colors = c("blue", "yellow"),type="surface", 
  #      hoverinfo="none", opacity=0.7, 
  #      legendshow=FALSE, showscale = FALSE) %>%
  layout(title = "Curved Kime Manifold")
p
```